\documentclass[final,1p]{elsarticle}
\usepackage{amssymb}
\usepackage{stackrel}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{float}
\usepackage{rotating}
\usepackage{color}
%\usepackage{ulem}
\usepackage{tikz}
\usepackage{epic}
% \usepackage{caption,subcaption}
%\usepackage{subeqnarray}
 

\newcommand\encircle[1]{%
  \tikz[baseline=(X.base)] 
    \node (X) [draw, shape=circle, inner sep=0] {\strut #1};}

\newcommand{\I}{\mathrm{i}}

\newcommand{\PGR}[1]{{\color{blue} #1}}
\newcommand{\PGRcomm}[1]{{\color{blue}\small\em PGR2all: #1}}
\newcommand{\PGRfoot}[1]{{\color{blue}\footnote{\color{blue} #1}}}
%\newcommand{\ES}[1]{{\color{red} #1}}
%\newcommand{\ESfoot}[1]{{\color{red}\footnote{\color{red} #1}}}
\newcommand{\bmu}{{\mbox{\boldmath{$\mu$}}}}

\newcounter{denselistcounter}
\newenvironment{denselist}[1]
{ \begin{list}
  {#1{denselistcounter})}{\usecounter{denselistcounter}
  \setlength{\topsep}{-0pt}
  \setlength{\partopsep}{-0pt}
  \setlength{\itemsep}{-0pt}
  \setlength{\parsep}{-0pt}
  \setlength{\labelwidth}{6pt}
  \setlength{\labelsep}{4pt}
  \setlength{\leftmargin}{20pt}
  \setlength{\rightmargin}{20pt}
  }
}
{\end{list}}

\unitlength 1mm
\thicklines

\begin{document}


\begin{frontmatter}

\title{Documentationm for the RTA code in 3D}

\author{F.~Coppens$^a$}
\author{M.~Vincendon$^a$}
\author{P.-G.~Reinhard$^c$}
\author{E.~Suraud$^{a,b}$}
\cortext[author]{Corresponding author:
  coppens@irsamc.ups-tlse.fr} 
\address{$^a$Universit\'e de Toulouse; UPS; Laboratoire de Physique
             Th\'{e}orique, IRSAMC; F-31062 Toulouse Cedex, France}
\address{$^b$Laboratoire de Physique Th\'eorique, Universit\'e Paul
  Sabatier, CNRS, F-31062 Toulouse C\'edex, France}
\address{$^c$Institut f{\"u}r Theoretische Physik, Universit{\"a}t
  Erlangen, D-91058 Erlangen, Germany}

\date{Status: 12. January 2019}
\begin{abstract}
\PGR{\em This is a first draft for the CPC presenting the TDLDA+RTA code
  to the public. The layout of presentation has yet to be discussed.
Presently it mixes theory and algorithm with code. We may also
consider collecting all theory \& numerics together then followed by a
huge block detailing the code.}
\end{abstract}

\begin{keyword}
%\PACS{05.30.Fk,31.70.Hq,34.10.+x,36.40.Cg}
electronic disspation, time-dependent
density functional theory, ...
\end{keyword}
\end{frontmatter}

\newpage



\section{Time-dependent local density approximation (TDLDA)}
\label{sec:TDLDA}

\PGRcomm{The main part of the code is very complex and much has to be
  explained. We need to address: LDA, pseudo-potentials, TDLDA, and
  MD. The question is how we arrange that in proper sectioning or
  whether we put generally known ``basics'' to appendices.
\\
The code is still capable of coupling to a dielectric
  environment. We should skip that for the CPC publication.
\\
Another question to be discussed is whether we maintain the option to
run finite differences instead of FFT. If we do so, then we have to
test that branch carefully.
}

\subsection{Mean-field propagation}
\label{sec:mf}

\PGRcomm{This piece is copied from our basic RTA paper and servers her
just to provide formulae needed later in RTA. It has yet to be
rewritten in connection with the new section on TDLDA-MD.}


The starting point and dominant feature of the dynamics is the
propagation at the level of the mean field. In this paper, we are
dealing with the electron dynamics in metal clusters and we describe
it by time-dependent density functional theory at the level of the
Time-Dependent Local-Density Approximation (TDLDA) treated in the real
time domain \cite{Gro90,Gro96}.  It is augmented by a self-interaction
correction (SIC) approximated by average-density SIC (ADSIC)
\cite{Leg02} in order to attain correct ionization properties
\cite{Klu13} in the course of the dynamical simulation. TDLDA is
formulated within the usual Kohn-Sham picture in terms of a set of
occupied single-particle (s.p.) wavefunctions
$\{|\phi_\alpha\rangle,\alpha=1...N\}$. Their dynamics is described by
the time-dependent Kohn-Sham equation
\begin{equation}
  \I\partial_t|\phi_\alpha\rangle
  =
  \hat{h}[\varrho]|\phi_\alpha\rangle
\label{eq:KSwf}
\end{equation}
where $\hat{h}$ is the Kohn-Sham mean-field Hamiltonian which is a
functional of the instantaneous local density
$\varrho(\mathbf{r},t)=\sum_\alpha|\phi_\alpha(\mathbf{r},t)|^2$
\cite{Rei04aB,Dre90}. The time evolution delivered by
Eq. (\ref{eq:KSwf}) can be expressed formally by the
unitary one-body  time-evolution operator 
\begin{subequations}
\begin{equation}
  \hat{U}(t,t')
  =
  \hat{\mathcal{T}}\mathrm{ exp}\left(-i \int_t^{t'} \hat{h}(t'')dt''\right)
\label{eq:KSpropag}
\end{equation}
where $\hat{\mathcal{T}}$ is the time-ordering operator.
This yields a closed expression for the time-evolution of s.p. states
\begin{equation}
  |\phi_\alpha(t)\rangle
  =
  \hat{U}(t,t')|\phi_\alpha(t')\rangle.
\label{eq:KSpropag2}
\end{equation}
\end{subequations}

So far, TDLDA propagates pure states. Dissipation which we will add
later on leads inevitably to mixed states. This requires to generalize
the description from fully occupied s.p. wavefunctions to a one-body
density operator $\hat{\rho}$. Its representation in
  configuration space, i.e. in terms of a given set of s.p. states
  $|\varphi_i\rangle$, reads in general
  $\hat{\rho}=\sum_{ij}|\varphi_i\rangle\rho_{ij}\langle\varphi_j|$.
  By appropriate transformation of the s.p. basis, one can diagonalize
  the density matrix $\rho_{ij}$ which defines what are called natural
  orbitals. The natural orbitals representation of the one-body
  density operator then reads
\begin{equation}
  \hat{\rho}
  =
  \sum_{\alpha=1}^\infty|\phi_\alpha\rangle W_\alpha\langle\phi_\alpha|
  \quad.
\label{eq:rhodiag}
\end{equation}
The weights $W_\alpha$ represent the probability with which a
state $|\phi_\alpha\rangle$ is occupied. The mean-field propagation
(\ref{eq:KSwf}) then becomes
\begin{equation}
  \I\partial_t\hat{\rho}
  =
  \left[\hat{h}[\varrho],\hat{\rho}\right]
\label{eq:KSrho}
\end{equation}
where $\hat{h}[\varrho]$ is formally the same as before and the
local density is now computed as
\begin{equation}
\varrho(\mathbf{r},t)=\sum_\alpha{W}_\alpha|\phi_\alpha(\mathbf{r},t)|^2 .
\label{eq:rhoW}
\end{equation}
The (coherent) pure mean-field propagation (\ref{eq:KSrho})
leaves the occupation weights $W_\alpha$ unchanged and propagates only
the s.p. states.  The mean-field propagation of an initial state
(\ref{eq:rhodiag}) then reads
\begin{eqnarray}
  \hat{\rho}(t)
  &=&
  \sum_{\alpha=1}^\infty
  |\phi_\alpha(t)\rangle W_\alpha\langle\phi_\alpha(t)|
%\nonumber\\  
%  &=&
  =
  \hat{U}(t,0)\hat{\rho}(0)\hat{U}^{-1}(t,0)
\label{eq:KSrhoevol}
\end{eqnarray}
where $\hat{U}$ is the mean-field evolution operator (\ref{eq:KSpropag}).


\section{The structure of the TDLDA package}
\label{sec:TDLDAnum}


\subsection{The TDLDA calling tree}

\PGRcomm{Here comes a calling tree as for RTA in section
  \ref{eq:treeRTA}. This tree is much larger and may have to be
  splitted into static and dynamic part.
}

\subsection{The TDLDA subroutines in detail}
\label{eq:detailsTDLDA}


\PGRcomm{Here is to come a detailed description of all subroutines
  similar as for RTA in \ref{eq:details}, but here for the LDA part.
One example is given as appetizer.
}

\subsubsection*{\tt SUBROUTINE coul\_mfield(rho)}
\begin{tabular}{lcl}
 {\tt rho(1:2*kdfull2)} & in/out & \\ density for which Coulomb field
 is computed
\end{tabular}
\\[4pt]
Computes Coulomb potential for given density by invoking
Poisson solver.
Th emerging coupomb potential is communicated as {\tt chpcoul}
via module {\tt params}.
In case of dielectric external media, adds pseudo-density for image
charge.
\\
\PGRcomm{Why is {\tt rho} also {\tt INTENT OUT}? Is all {\tt
    1:2*kdfull2} used or only the first block {\tt 1:kdfull2}?}



%\subsubsection*{\tt }
%\begin{tabular}{lcl}
% {\tt } & &\\
%\end{tabular}
%\\[4pt]




\section{Relaxation-time approximation (RTA)}


\subsection{The formal background of RTA}

The quantum Boltzmann equation is the quantum mechanical
counterpart of the
semiclassical Vlasov-Uehling-Uhlenbeck equation \cite{Ber88,Abe96}.
It complements the self-consistent TDLDA propagation 
of the one-body density matrix $\hat{\rho}$ by
dynamical correlations through a collision term. It reads in
general \cite{Rei85f,Goe86a}
%\begin{eqnarray}
$\mathrm{i}\partial_t\hat{\rho}
  -
  \big[\hat{h},\hat{\rho}\big]
  =
  \hat{I}[\hat{\rho}]
%  \quad.
$
%\label{eq:EoMfull}
%\end{eqnarray}
where the left hand side contains the mean-field propagation. the
$\hat{I}$ at the right-hand side consists stands for the
quantum-mechanical collision term which, however, is extremely hard to
handle for finite Fermion systems. A great simplification can be
achieved by the relaxation-time approximation (RTA) which was used
successfully in a wide variety of homogeneus systems
\cite{Pin66,Ash76}. The RTA equations for the present case of finite
Fermion systems read \cite{Rei15a}
\begin{subequations}
\label{eq:EoMbasic}
\begin{eqnarray}
  \partial_t\hat{\rho}
  &=&
  -\mathrm{i}\big[\hat{h},\hat{\rho}\big]
  -
  \frac{1}{\tau_\mathrm{relax}}
  \left(\hat{\rho}-\hat{\rho}_\mathrm{eq}[\varrho,\mathbf{j},E]\right)
  \;,
\label{eq:EoMbasicrho}
\\
  \varrho(\mathbf{r},t)
  &=&
  \sum_\alpha \left|\phi_\alpha(\mathbf{r},t)\right|^2 W_\alpha
  \;,
\\
  \mathbf{j}(\mathbf{r},t)
  &=&
  \sum_\alpha W_\alpha\phi_\alpha^*(\mathbf{r},t)
     \frac{\stackrel{\rightarrow}{\nabla}-\stackrel{\leftarrow}{\nabla}}
          {2\mathrm{i}}
     \phi_\alpha(\mathbf{r})
  \;,
\label{eq:locdens}\\
  \frac{\hbar}{\tau_\mathrm{relax}}
  &=&
  {0.40}\frac{\sigma_{ee}}{r_s^2}\frac{{E}^*_\mathrm{intr}}{N}
  \;,\;
  r_s=\left(\frac{3\bar{\varrho}}{4\pi}\right)^{-1/3}
  \;,\;
  \sigma_{ee}=\sigma_{ee}(\bar{\varrho})
  \;,
\label{eq:relaxtime}
\end{eqnarray}
\end{subequations}
where $\hat{\rho}_\mathrm{eq}$ is the density operator of the thermal
equilibrium for local density $\varrho(\mathbf{r},t)$, current
distribution $\mathbf{j}(\mathbf{r},t)$ and total energy $E(t)$ given
at that instant of time $t$ and computed from the actual state
$\hat{\rho}(t)$.  The form (\ref{eq:locdens}) holds for the diagonal
representation.  
A crucial parameter is the relaxation time
$\tau_\mathrm{relax}$ which is taken over from semi-classical Fermi
liquid theory, for details see \cite{Rei15a}.  Key entries are: the
intrinsic (thermal) energy of the system $E^*_\mathrm{intr}$ (see
appendix \ref{app:eintr}), the actual number of particles $N$, the
in-medium electron-electron cross section $\sigma_{ee}$, the effective
Wigner-Seitz radius $r_s$ of the electron cloud, and the average
electron density $\bar{\varrho}$.  Note that $r_s$ and $\sigma_{ee}$
depend on an average density $\bar{\varrho}$ because a spatially
varying $\tau_\mathrm{relax}$ would be very cumbersome to implement in
a quantum mechanical expression.  The average density is deduced from
the r.m.s. radius $r$ of the actual electron cloud as 
$\bar{\varrho}=(3N/(4\pi r^3))$\PGRfoot{Should we add that as
  displayed equation?}.

This RTA equation (\ref{eq:EoMbasic}) is rather involved because its
entries depend in various ways on the actual state $\hat{\rho}(t)$.
The most expensive piece is the instantaneous equilibrium density
$\hat{\rho}_\mathrm{eq}$ which minimizes LDA energy with constraint on
the actual $\varrho(\mathbf{r})$, $\mathbf{j}(\mathbf{r})$ and energy
$E$.  It is determined by the density-constrained mean-field (DCMF)
equation
\begin{subequations}
\label{eq:DCMF}
\begin{eqnarray}
  &&
  \left(
  \hat{h}
  -
  \int d^3r\lambda(\mathbf{r})\hat{\varrho}(\mathbf{r})
  -
  \int d^3r\,\mbox{\boldmath$\lambda_j$}(\mathbf{r})\hat{\mathbf{j}}(\mathbf{r})
  \right)
  \phi_\alpha^\mathrm{(eq)}
  =
  \varepsilon_\alpha^\mathrm{(eq)}\phi_\alpha^\mathrm{(eq)}
\\
  &&
  W_\alpha^\mathrm{(eq)}
  =
  \frac{1}{1+
  \exp((\langle\phi_\alpha^\mathrm{(eq)}|\hat{h}|\phi_\alpha^\mathrm{(eq)}\rangle-\mu^\mathrm{(eq)})/T^\mathrm{(eq)})}
\end{eqnarray}
\end{subequations}
where the chemical potential $\mu^\mathrm{(eq)}$ is adjusted to the
given particle number by $N(t)=\sum_\alpha W_\alpha^\mathrm{(eq)}$ and
the temperature $T^\mathrm{(eq)}$ is tuned such the the LDA energy
reproduces the wanted energy $E(t)$.
%
Although cumbersome to evaluate, it is important to use exactly this
local, instantaneous equilibrium in the relaxation term. This
guarantees that the disspative step conserves local density, 
current, and energy as it is mandatory for a good collision term
\cite{Gue88a}.


\subsection{Observables specific to relaxation}

Most of the observables computed with RTA are exactly the same as for
TDLDA, e.g., energy, density, excitation spectra, ionization, PES, or PAD.
New are observables related to the mixed character of the
one-body operator which is characterized by the occupation
numbers $W_\alpha$. 
A specific quantity in that respect is the 
entropy which is computed in diagonal representation
(\ref{eq:rhodiag}) by the standard expression \cite{Rei98aB}
\begin{equation}
  S
  =
  - \sum_\alpha\left[
    W_\alpha\log W_\alpha
    +
    (1\!-\!W_\alpha)\log (1\!-\!W_\alpha)
  \right]
\label{eq:entropy}
\end{equation}
in units of Boltzmann constant. 
It serves as a direct indicator of thermalization and allows to 
read off the typical time scale of relaxation processes. 



\subsection{Summary of the RTA procedure}
\label{sec:summaryRTA}




\begin{figure}
%\setlength\unitlength{0.91mm}
\thicklines
\begin{center}
%\begin{sideways}
\fbox{\small%\footnotesize
%\begin{picture}(130,92)(0,108)
\begin{picture}(132,175)(-5,-175)
\put(0,-5){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
$\;$starting point: 
\\[4pt]
  $\hat{\rho}(t_0)
  =
  \sum_\alpha|\phi_\alpha(t_0)\rangle W_\alpha(t_0)\langle\phi_\alpha(t_0)|$
\end{flushleft}
\end{minipage}
}}
\put(5,-19){\encircle{1}}
\put(3,-11.5){\vector(0,-1){25}}
\put(6,-30){\fbox{\tt tstep}}
\put(7,-25){\mbox{mean-field propagation: 
$%\begin{array}{l}
 |\phi_\alpha^\mathrm{(mf)}\rangle=\hat{U}(t_1,t_0)|\phi_\alpha(t)\rangle
\;,\;
%\\
 W_\alpha^\mathrm{(mf)}=W_\alpha(t)=\mbox{const.}
%\end{array}
$
}}
\put(0,-40){\mbox{
  $ \hat{\rho}_\mathrm{mf} =\hat{\rho}_\mathrm{mf}(t_1) 
  =
  \sum_\alpha|\phi_\alpha^\mathrm{(mf)}\rangle 
  W_\alpha^\mathrm{(mf)}\langle\phi_\alpha^\mathrm{(mf)}|$
}}
\put(5,-46){\encircle{2}}
\put(6,-54.5){\fbox{\tt srhomat}}
\put(3,-41.5){\vector(0,-1){15}}
\put(0,-50){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
\hspace*{1.8em}express $\phi$ and $W$ through natural orbitals
\\[20pt]
  $\displaystyle
  \hat{\rho}_\mathrm{mf}(t_1)
  =
  \sum_\alpha|\phi_\alpha^\mathrm{nat}\rangle \tilde{W}_\alpha^\mathrm{nat}
  \langle\phi_\alpha^\mathrm{nat}|$
\end{flushleft}
\end{minipage}
}}
\put(44,-63){\vector(4,-1){24}}
\put(51,-62){\encircle{3}}
\put(61,-64){\fbox{\tt calcrhotot,calc\_current}}
\put(69,-70){\mbox{  $\varrho_\mathrm{mf}(\mathbf{r},t_1)\,,\,\mathbf{j}_\mathrm{mf}(\mathbf{r},t_1),E_\mathrm{mf}$}}

\put(87,-72){\vector(1,-1){12}}
%\put(117,-103.5){\vector(0,-1){8}}
\dottedline{1}(117,-103.5)(117,-110.5)
\put(117,-110.5){\vector(0,-1){1}}
\put(85,-77){\encircle{4}}
\put(97,-79){\fbox{\tt eqstate}}
\put(67,-87){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
  density-constrained mean field (DCMF) %\\ $\hookrightarrow$ %$\Longrightarrow$
\\[3pt]
$\Rightarrow$ 1. 
$\begin{array}[t]{rcl}
  \hat{\rho}_{eq}
  &=&
  \hat{\rho}_{eq}[\varrho_\mathrm{mf}(\mathbf{r})\,,\,\mathbf{j}_\mathrm{mf}(\mathbf{r}),E_\mathrm{mf}]
\\[3pt]
  &=&
  \sum_\alpha|\phi'_\alpha\rangle  W'_\alpha\langle\phi'_\alpha|
\end{array}$
\\[4pt]
$\Rightarrow$ 2. intrinsic excitation energy $E^*_\mathrm{intr}$
\end{flushleft}
\end{minipage}
}}
%\put(79,-93.5){\vector(-4,-1){33}}
\dottedline{1}(77,-93.5)(47,-101.5)
\put(47,-101.5){\vector(-4,-1){1}}
\put(83,-110){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
relaxation time:
\\[4pt]
$\hbar\tau_\mathrm{relax}^{-1}=0.40\,\sigma_{ee}r_s^{-2}\,E^*_\mathrm{intr}/N$
\end{flushleft}
\end{minipage}
}}
\dottedline{1}(83,-114)(31,-106)
\put(31,-106){\vector(-4,1){1}}
\put(4,177){\vector(0,-1){39}}
\put(0,-96){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
\hspace*{1.8em}compose with rate $\tau_\mathrm{relax}$:
\\[4pt]
$\displaystyle
\hat{\rho}_\mathrm{mix}
= 
\hat{\rho}_\mathrm{mf} -
\frac{\Delta t}{\tau_\mathrm{relax}}\left[\hat{\rho}_\mathrm{mf}-\hat{\rho}_{eq}\right]
$
\end{flushleft}
\end{minipage}
}}
\put(3,-104){\vector(0,-1){31}}
\put(5,-121){\encircle{5}}
\put(6,-130.5){\fbox{\tt calcrhoeq}}
\put(0,-126){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
\hspace*{1.8em}express new density through its natural orbitals:
\\[24pt]
  $\displaystyle
  \hat{\rho}_\mathrm{mix}
  =
  \sum_\alpha|\phi_\alpha(t_1)\rangle \tilde{W}_\alpha
  \langle\phi_\alpha(t_1)|$
\end{flushleft}
\end{minipage}
}}
\put(3,-140){\vector(0,-1){21}}
\put(5,-146.5){\encircle{6}}
\put(6,-155.5){\fbox{\tt temperature}}
\put(0,-151){\mbox{
\begin{minipage}[t]{8cm}
\begin{flushleft}
\hspace*{1.8em}final fine-tuning of $W_\alpha$ to reproduce $E_\mathrm{mf}$
\\[24pt]
  $\displaystyle
  \hat{\rho}(t_1)
  =
  \sum_\alpha|\phi_\alpha(t_1)\rangle W_\alpha(t_1)\langle\phi_\alpha(t_1)|$
\end{flushleft}
\end{minipage}
}}
\dottedline{1}(3,-166)(3,-171)(-1.5,-171)(-1.5,-4.2)(0,-4.2)
\put(0,-4.2){\vector(1,0){1}}
\end{picture}
}
%\end{sideways}
\end{center}
\caption{\label{fig:summary} Sketch of the scheme for performing one
  large time step $t_0\longrightarrow t_1=t_0\!+\!\Delta t$ in solving
  the RTA equations.  The numbers in open circles indicate the steps
  as outlined in the text. The names in {\tt typewriter} font refer to
  subroutines in the code as detailed in section \ref{eq:RTApack}. }
\end{figure}



The solution of the RTA equations (\ref{eq:EoMbasic}) with
(\ref{eq:DCMF}) is rather involved. We briefly summarize the solution
scheme for one step from $t\equiv t_0$ to $t\!+\!\Delta t\equiv t_1$,
for more details see \cite{Rei15a}\PGRfoot{PGR2all: can we outsource
  all details to reference \cite{Rei15a}?}.
The TDLDA propagation runs at a much faster pace than relaxation.  We
resolve it by standard techniques \cite{Cal00,Rei04aB} on a time step
$\delta t$ which is much smaller (factor 10--100) than the RTA step
$\Delta t$. We summarize this TDLDA propagation in the evolution
operator $\hat{U}$ from Eq.~(\ref{eq:KSpropag}) and discuss only one
RTA step. Its sub-steps are sketched in Figure \ref{fig:summary} and
explained in the following whereby the step numbers here correspond to
the ones in the Figure:
\begin{enumerate}
   \item\label{it:TDLDA} We first propagate $\hat{\rho}$ by pure
     TDLDA.  The s.p. states in diagonal representation
     (\ref{eq:rhodiag}) evolve as $|\phi_\alpha(t)\rangle\rightarrow
     |\phi_\alpha^\mathrm{(mf)}\rangle=\hat{U}(t_1,t_0)|\phi_\alpha(t)\rangle$,
     while the occupation weights $W_\alpha(t_1)=W_\alpha(t_0)$ are
     kept frozen (pure mean-field propagation).
   \item\label{it:natorb} Absorbing bounds may have removed parts
     from the s.p. wavefunctions and so destroyed ortho-normalization.
     We transform the propagated density operator to a representation
     in terms of natural orbitals which is  diagonal representation
     (\ref{eq:rhodiag}) with an ortho-normal set of s.p. wavefunctions
     together with corresponding occupation weights
     $\{\phi_\alpha^\mathrm{(nat)},W_\alpha^\mathrm{(nat)}\}$.
   \item\label{it:newrho} We compute density
     $\varrho(\mathbf{r},t_1)$, current
     $\mathbf{j}(\mathbf{r},t_1)$, and total energy
     $E_\mathrm{mf}$ associated to the TDLDA-propagated density matrix
     $\hat{\rho}_\mathrm{mf}$.
   \item\label{it:DCMF} We determine the thermal mean-field
     equilibrium state $\hat{\rho}_\mathrm{eq}$ constrained to the
     given $\varrho$, $\mathbf{j}$, and $E_\mathrm{mf}$ from step
     \ref{it:newrho}.  This is achieved by Density-Constrained Mean
     Field (DCMF) iterations as outlined in section
     \ref{sec:summaryDCMF}.  The equilibrium state
     $\hat{\rho}_\mathrm{eq}$ is represented by new s.p. states
     $\{|\phi'_{\alpha}\rangle\}$ and new occupation numbers
     $W'_\alpha$ in diagonal form (\ref{eq:rhodiag}).
   \item \label{it:compo} We compose the new density operator as mixture
     of TDLDA propagated state $\hat{\rho}_\mathrm{mf}$ and
     equilibration driving term
     $\hat{\rho}_\mathrm{mf}-\hat{\rho}_\mathrm{eq}$ with
     weight $\Delta t/\tau_\mathrm{relax}$ as
     \begin{equation*}
        \hat{\rho}_\mathrm{mix}
        = 
        \hat{\rho}_\mathrm{mf} -
        \frac{\Delta t}{\tau_\mathrm{relax}}
        \left[\hat{\rho}_\mathrm{mf}-\hat{\rho}_{eq}\right]
     \end{equation*}
     where the relaxation time $\tau_\mathrm{relax}$ requires the
     actual intrinsic excitation energy $E^*_\mathrm{intr}$ which is
     also obtained from DCMF.
%   \item \label{it:natural} 
     While evaluating the mixing, the new state is
     expressed in natural-orbital representation
     Eq. (\ref{eq:rhodiag}).  This yields the final s.p. states
     $\{|\phi_\alpha(t_1)\rangle\}$ for this step and
     preliminary new occupations $\tilde{W}_\alpha$.
   \item \label{it:therm} 
     The mixing in step \ref{it:compo} may have slightly changed
     the energy such that we remain with a small energy mismatch as
     compared to the goal $E_\mathrm{mf}$.  We now apply a small
     iterative thermalization step to readjust the energy, as outlined
     in Appendix \ref{sec:corriter}. This then yields the final
     occupation weights $W_\alpha(t_1)$ which comply with
     energy conservation.
\end{enumerate}
The scheme is designed for use in connection with absorbing boundary
conditions \cite{Cal00,Rei06c}. 
In case of reflecting boundaries, natural orbital property is
maintained under TDLDA propagation and step \ref{it:natorb}
can be overridden.

As said above, the time step $\delta t$ for propagation of TDLDA is
very small because it is limited from above by the maximal energy on
the grid representation.  The stepping $\Delta t$ for the relaxation
term needs only to resolve the changes in the actual mean field which
allows much larger values.  Typically, we can do 50--200 TDLDA steps
before calling one RTA step. For detailed values see the examples
delivered with the code.\PGRfoot{PGR2all: A link to be set once we
have the place for the benchmarks.}


A word is in order about the system for which the present form of RTA
can be used.  The relaxation time $\tau_\mathrm{relax}$ is allowed to
depend on time which allows to accomodate changes of the dynamical
state. But $\tau_\mathrm{relax}$ is one global number chosen according
to the average electron density $\bar{\varrho}$, see
eq. (\ref{eq:relaxtime}).  This requires systems which can be
characterized by such an average density, i.e., systems having only
small density variations in the bulk as it holds typically for
metallic bonds.  The RTA rate is insensitive to many details of the
microscopic collision term as energy- and angle-dependent scattering
cross sections \cite{Gig03a} or a broad spectrum of relaxation
rates. However, these details are usually resolved only (if at all)
for fast and energetic processes which are anyway deep in the regime
of semi-classical VUU. The grossly averaged treatment of RTA is
acceptable for not too fast and not too energetic processes,
preferably in compact systems.


\section{Summary of the DCMF procedure}
\label{sec:summaryDCMF}


\PGRcomm{This section for DCMF in the same layout as for RTA, i.e. a
  summary in text together with an even compacter summary diagram.}

\section{The structure of the RTA package in {\tt rta.F90}}
\label{eq:RTApack}

\subsection{The calling tree}
\label{eq:treeRTA}

Here is an oversight over the tree structure of the RTA routines.
Those subroutines contained in  {\tt rta.F90} are explained in detail
in section \ref{eq:details}. Subroutines coming from the TDLDA package
or external sources are marked\PGRfoot{The {\tt HEigensystem} seems
  copied from some library. This could cause copyright problems if we
  publish the code. Is it from BLAS/LINPACK? Then we could
replace the Fortran source by a library call.}
\\
\centerline{\fbox{
\begin{picture}(122,159)(5,-155)
\put(10,0){\mbox{\tt RTA}}
\put(11,-1){\line(0,-1){150}}
   \put(20,-4){\mbox{\tt srhomat}}
   \put(11,-3){\line(1,0){8}}
   \put(21,-5){\line(0,-1){6}}
      \put(30,-8){\mbox{\tt scalar}}
      \put(21,-7){\line(1,0){8}}
      \put(30,-12){\mbox{\tt cdiagmat}}
      \put(21,-11){\line(1,0){8}}
      \put(31,-13){\line(0,-1){2}}
         \put(40,-16){\mbox{\tt HEigensystem}}
         \put(31,-15){\line(1,0){8}}
             \put(90,-16){\mbox{$\leftrightarrow$ library routine}}
   \put(20,-20){\mbox{\tt eqstate}}
   \put(21,-21){\line(0,-1){74}}
   \put(11,-19){\line(1,0){8}}
      \put(30,-24){\mbox{\tt calcrhotot}}
      \put(21,-23){\line(1,0){8}}
      \put(30,-28){\mbox{\tt calc\_current}}
      \put(21,-27){\line(1,0){8}}
      \put(30,-32){\mbox{\tt calc\_Eref}}
      \put(21,-31){\line(1,0){8}}
      \put(30,-36){\mbox{\tt fermi1}}
      \put(31,-37){\line(0,-1){2}}
      \put(21,-35){\line(1,0){8}}
         \put(40,-40){\mbox{\tt occT1}}
         \put(31,-39){\line(1,0){8}}
      \put(30,-44){\mbox{\tt calc\_psi1}}
      \put(31,-45){\line(0,-1){38}}
      \put(21,-43){\line(1,0){8}}
         \put(40,-48){\mbox{\tt calc\_hamiltonien}}
         \put(31,-47){\line(1,0){8}}
         \put(40,-52){\mbox{\tt calc\_var}}
         \put(41,-53){\line(0,-1){2}}
            \put(50,-56){\mbox{\tt cproject}}
            \put(41,-55){\line(1,0){8}}
            \put(90,-56){\mbox{\PGR{$\leftrightarrow$ \footnotesize obsolete?}}}
         \put(40,-60){\mbox{\tt calcrhotot}}
         \put(31,-59){\line(1,0){8}}
         \put(40,-64){\mbox{\tt calc\_current}}
         \put(31,-63){\line(1,0){8}}
         \put(40,-68){\mbox{\tt cschmidt}}
         \put(31,-67){\line(1,0){8}}
             \put(90,-68){\mbox{$\leftrightarrow$ TDLDA package}}
         \put(40,-72){\mbox{\tt calc\_ekin}}
         \put(31,-71){\line(1,0){8}}
             \put(90,-72){\mbox{$\leftrightarrow$ TDLDA package}}
         \put(40,-76){\mbox{\tt nonlocalc}}
         \put(31,-75){\line(1,0){8}}
             \put(90,-76){\mbox{$\leftrightarrow$ TDLDA package}}
         \put(40,-80){\mbox{\tt fftf}}
         \put(31,-79){\line(1,0){8}}
             \put(90,-80){\mbox{$\leftrightarrow$ FFTW3 package}}
         \put(40,-84){\mbox{\tt fftback}}
         \put(31,-83){\line(1,0){8}}
             \put(90,-84){\mbox{$\leftrightarrow$ FFTW3 package}}
      \put(30,-88){\mbox{\tt coul\_mfield}}
      \put(21,-87){\line(1,0){8}}
             \put(90,-88){\mbox{$\leftrightarrow$ TDLDA package}}
      \put(30,-92){\mbox{\tt dyn\_mfield}}
      \put(21,-91){\line(1,0){8}}
             \put(90,-92){\mbox{$\leftrightarrow$ TDLDA package}}
      \put(30,-96){\mbox{\tt info}}
      \put(21,-95){\line(1,0){8}}
             \put(90,-96){\mbox{$\leftrightarrow$ TDLDA package}}
   \put(20,-100){\mbox{\tt occupT0}}
   \put(21,-101){\line(0,-1){2}}
   \put(11,-99){\line(1,0){8}}
      \put(30,-104){\mbox{\tt indexx}}
      \put(21,-103){\line(1,0){8}}
   \put(20,-108){\mbox{\tt calcrhoeq}}
   \put(21,-109){\line(0,-1){10}}
   \put(11,-107){\line(1,0){8}}
      \put(30,-112){\mbox{\tt cdiagmat}}
      \put(31,-113){\line(0,-1){2}}
      \put(21,-111){\line(1,0){8}}
         \put(40,-116){\mbox{\tt HEigensystem}}
         \put(31,-115){\line(1,0){8}}
             \put(90,-116){\mbox{$\leftrightarrow$ library routine}}
      \put(30,-120){\mbox{\tt  indexx}}
      \put(21,-119){\line(1,0){8}}
   \put(20,-124){\mbox{\tt calc\_Eref}}
   \put(11,-123){\line(1,0){8}}
   \put(20,-128){\mbox{\tt CorrectEnergy2}}
   \put(11,-127){\line(1,0){8}}
   \put(20,-132){\mbox{\tt OccupPerSpin}}
   \put(11,-131){\line(1,0){8}}
   \put(20,-136){\mbox{\tt temperature}}
   \put(11,-135){\line(1,0){8}}
   \put(21,-137){\line(0,-1){2}}
      \put(30,-140){\mbox{\tt lmdif1}}
      \put(21,-139){\line(1,0){8}}
   \put(20,-144){\mbox{\tt dyn\_mfield}}
   \put(11,-143){\line(1,0){8}}
             \put(90,-144){\mbox{$\leftrightarrow$ TDLDA package}}
   \put(20,-148){\mbox{\tt info}}
   \put(11,-147){\line(1,0){8}}
             \put(90,-148){\mbox{$\leftrightarrow$ TDLDA package}}
   \put(20,-152){\mbox{\tt analyze\_elect}}
   \put(11,-151){\line(1,0){8}}
             \put(90,-152){\mbox{$\leftrightarrow$ TDLDA package}}
\end{picture}
}}

\subsection{The subroutines in detail}
\label{eq:details}

\subsubsection*{\tt SUBROUTINE rta(psi,aloc,rho,iterat)}
\begin{tabular}{lcl}
 {\tt iterat} & in & external iteration number (TDLDA time step)\\
 {\tt psi(1:kdfull2,1:kstate)} & in/out& set of s.p. wavefunctions\\
 {\tt rho(1:2*kdfull2)}& in/out & local densities for spin up and down\\
 {\tt aloc(1:2*kdfull2)} & in/out& local potentials for spin up and down\\
\end{tabular}
\\[4pt]
Basic RTA routine performing density constrained mean-field (DCMF)
iterations, energy adjustment, admixing of local equilibrium states by
calls to subroutines (see calling tree).



\subsubsection*{\tt SUBROUTINE calcrhoeq(psiorthloc,psieqloc,psiloc,occuporthloc,occuploc,nstateloc)}
\begin{tabular}{lcl}
 {\tt nstateloc} & in & number of s.p. states in spin block\\
 {\tt psiorthloc(1:kdfull2,1:nstateloc),} & in & set of TDLDA
 wavefunctions (natural orbitals)\\
 {\tt occuporthloc(1:nstateloc)} & in & occupations of TDLDA states\\
 {\tt psieqloc(1:kdfull2,1:nstateloc)} & in& set of local-equilibrium wavefunctions\\
 {\tt psiloc(1:kdfull2,1:nstateloc)} & out& set of final mixed wavefunctions\\
 {\tt occuploc(1:nstateloc)} & in/out & \\
\end{tabular}
\\[4pt]
Encapsulated in {\tt SUBROUTINE rta}. Performs the mixing of TDLDA
states with local-equilibrium state according to relaxation rate
for one spin block. The mixed densty matrix is expanded in a
representation by both sets of s.p. states. 



\subsubsection*{\tt SUBROUTINE calc\_Eref(occup,ispin,Ei,Eref)}
\begin{tabular}{lcl}
 {\tt occup(1:nstate)} & in & occupation number for s.p. states.\\
 {\tt ispin(1:nstate)} & in & spin assignement for s.p. states.\\
 {\tt Ei(1:nstate)} & in & spin assignement for s.p. states.\\
 {\tt Eref(1:2)} & out & sum of s.p. energies per spin.\\
\end{tabular}
\\
Computes the weighted sum of s.p. energies as reference energy
for DCMF. The sum is accumulated for each spin separately.


\subsubsection*{\tt SUBROUTINE fermi1(ekmod,eref,occup,ispinact,T0i,T1i,T2,mu)}
\begin{tabular}{lcl}
 {\tt ekmod(1:kstate)} & in & given s.p. energies, spin up block first, then
 spin down\\
 {\tt eref}& in & reference energy = wanted sum of s.p. energies\\
 {\tt ispinact}& in & spin for which routine is run\\
 {\tt T0i, T1i}& in & lower and upper temperature for search\\
 {\tt occup(1:kstate)}& in/out & occupation numbers, spin block-wise\\
 {\tt T2} & out & final temperature for which Fermi distribution
 matches {\tt eref} \\
 {\tt mu} & out & final chemical potential\\
\end{tabular}
\\[4pt]
Determines thermal Fermi occupation such that given sum of
s.p. energies {\tt eref} and particle number is matched. Is done for
each spin separately. Solution scheme is bracketing. Refers to 
{\tt SUBROUTINE OccT1} while iterating temperatur {\tt T2}.
\\
\PGRcomm{Nr. of spin-up/spin-down states comes through {\tt m\_params}. 
We should protocol all such entries. First step is to augment each
{\tt USE} by {\tt ONLY} such that the explicitely communicated
variables becomes visible. Important variables may then be listed
explicitely.
}
\\
\PGRcomm{Routine requires that arrays are sorted in continuous blocks of
  spin. Do we have an initial check for that? And we need to address
  that in the general part which explains the layout of arrays.}

\subsubsection*{\tt SUBROUTINE OccT1(occrefloc,enerloc,Etotloc,muloc,occtotloc,n,T,occuploc)}
\begin{tabular}{lcl}
 {\tt enerloc(1:n)} & in & s.p. energies for actual spin\\
 {\tt n} & in & number of s.p. states treated here\\
 {\tt T} & in & temperature\\
 {\tt occrefloc} & in & wanted total number of particles \\
 {\tt occuploc(1:n)} & out & thermal occupation numbers for given {\tt
   T} and s.p. energies\\
 {\tt muloc} & out & chemical potential (Fermi energy)\\
 {\tt occtotloc} & out& final total number of particles \\
 {\tt Etotloc} & out& sum of s.p. energies \\
\end{tabular}
\\[4pt]
Determines by bracketing chemical potential {\tt muloc} for given array of
s.p. energies, temperature {\tt T}, and wanted number of particles
{\tt occrefloc} with precision {\tt 1D-12}. Delivers with it
thermal occupation numbers and corresponding total particle number and
sum of s.p. energies.
\\
\PGRcomm{This routine is specific to {/tt SUBROUTINE ferm1}. Could we
encapsulate it by a {/tt CONTAINS}?}

\subsubsection*{\tt SUBROUTINE
  Calc\_psi1(psi1,aloc,rhotot0,rhototloc,curr0,curr1,j,lambda,mu,lambdaj,muj,sumvar2,eal,ekmod)}
\noindent
combined with encapsulated {\tt SUBROUTINE calc\_hamiltonien}.
\\
\begin{tabular}{lcl}
 {\tt j} & in & number of DCMF iteration, used here for print\\
 {\tt lambda(1:kdfull2,1:2)} & in & Lagrange parameter for density for
 spin up\&down\\
 {\tt lambdaj(1:kdfull2,1:3)} & in & Lagrange parameter for current\\
 {\tt mu, muj} & in & driving parameter for augmented Lagrangian\\
 {\tt aloc(1:2*kdfull2)} & in & local potentials for spin up and down\\
 {\tt rhoto0(1:kdfull2,1:2)} & in & initial density \PGRcomm{not used ??}\\
 {\tt curr0(1:kdfull2,1:3)} & in & wanted current \\
 {\tt psi1(1:kdfull2,1:kstate)} & in/out & set of s.p. wavefunctions iterated\\
 {\tt rhototloc(1:kdfull2,1:2)} & out & actual density according to {\tt psi1}\\
 {\tt curr1(1:kdfull2,1:3)} & out & actual current from {\tt psi1}\\
 {\tt ekmod(1:nstate)} & out & final s.p. energies\\
 {\tt eal} & out & final sum of s.p. energies\\
 {\tt sumvar2} & out & variance of s.p. energies\\
\end{tabular}
\\[4pt]
Performs one damped gradient step of with density \& current
constrained Hamiltonian.
\\
\PGRcomm{The density array distinguishes spin up/down while the
  current array does not. Reason?}
\\
PGRcomm{The IN \& OUT assignments in this subroutine have to be updated.}



\subsubsection*{\tt SUBROUTINE eqstate(psi,aloc,rho,psi1,occuporth,iterat)}
\begin{tabular}{lcl}
 {\tt iterat} & in & actual iteration number (for printing)\\
 {\tt psi(1:kdfull2,1:kstate)} & in & initial set of
 s.p. wavefunctions\\
 {\tt psi1(1:kdfull2,1:kstate)} & out & final set of s.p. wavefunctions\\
 {\tt aloc(1:2*kdfull2)} & in/out & local part of potential, spin
 up/down stacked in blocks\\
 {\tt rho(1:2*kdfull2)} & in & initial density, spin
 up/down stacked in blocks \\
 {\tt occuporth(1:kstate)} & in & occupation numbers for {\tt psi} and
 still the same for {\tt psi1}.\\
\end{tabular}
\\[4pt]
DCMF iterations by reapeatedly calling {\tt Calc\_psi1},
updating Lagrangian parameters for density \& current constraints, and
occassionally tuning temperature to achieve correct energy. The
latter is done by calling {\tt fermi1}. The local potential
is kept constant during DCMF iteration and updated only at the very end.
\\
\PGRcomm{Fetches nr. of spin up/down from {\tt m\_params}.}
\\
\PGRcomm{Lagrange parameters are started from scratch. May it be
  faster to recycle the previous Lagrange parameters?
}
\\
\PGRcomm{Density {\tt rho} is entered via list and still recomputed
as {\tt rhotot0}. Unnecessary doubling?}


\subsubsection*{\tt SUBROUTINE OccupT0(occloc,esploc,Estar)}
\begin{tabular}{lcl}
 {\tt esploc(1:nstate)} & in & given s.p. energies\\
 {\tt occloc(1:nstate)} & in & given occupation numbers\\
 {\tt Estar} & out & excitation energy relative to T=0 distribution\\
\end{tabular}
\\[4pt]
Computes thermal excitation energy as difference of actual energy to
the energy obtained by Fermi distribution for $T=0$. The latter
distributions  is computed for the given s.p. energies which are the
same as used for the thermal state.



\subsubsection*{\tt SUBROUTINE calcrhotot(rho,q0)}
\begin{tabular}{lcl}
 {\tt q0(1:kdfull2,1:kstate) } & in & set of s.p. wavefunctions for
 which density is accumulated\\
 {\tt rho(kdfull2,2)} & out & resulting density\\
\end{tabular}
\\[4pt]
Computes local density for set of wavefunctions {\tt q0}. Note that
two crucial information is communicated via module {\tt params}, namely
{\tt occup}, the array of occupation numbers, {\tt ispin} the
array assigning spin top each s.p. state, and {\tt nstate}, the number
of s.p. states.
\\
\PGRcomm{Exploiting the sorting of spin in blocks of s.p. states, we
  could rewrite the code with to {\tt SUM} statements.}


\subsubsection*{\tt SUBROUTINE calc\_var(hpsi,psi1,sumvar2)}
\begin{tabular}{lcl}
 {\tt psi1(kdfull2,kstate)}& in & set of s.p. states for which
 variance of s.p. energies of calculated\\
 {\tt hpsi(kdfull2,kstate)} & in/out & array
 $H\rightarrow\psi_\alpha$, on input in $k$-space, on output in $r$-space
\\
 {\tt sumvar2} & out  & summed variance of s.p. energies\\
\end{tabular}
\\[4pt]
Computes the sum of variances of the s.p. energies,
$\langle\hat{\Delta h}^2|rangle$. 
\\
PGRcomm{The routine projects from each $hat{h}\psi_\alpha$
all s.p. states $\psi_\beta$ from the pool of states. That is too
much. The s.p. variance should be
$\sum_\alpha\langle|\psi_\alpha|(\hat{h}-\varepsilon_\alpha)^2|\psi_\alpha\rangle$
where $\varepsilon_\alpha=\langle|\psi_\alpha|\hat{h}|\psi_\alpha\rangle$.}




\subsubsection*{\tt SUBROUTINE forceTemp(amoy,occup,n,temp,mu)}
\begin{tabular}{lcl}
 {\tt amoy(1:n)} & in & given s.p. energies\\
 {\tt occup(1:n)} & in & given thermal occupation\\
 {\tt n} & in & number of s.p. states\\
 {\tt temp} & in & temperature\\
 {\tt mu} & out & emerging chemical potential\\
\end{tabular}
\\[4pt]
Determines chemical potential for given s.p. energies and temperature
by call to {\tt OccT1}.
\\
\PGRcomm{Obsolete and never used.}




\subsubsection*{\tt SUBROUTINE fermi\_init(ekmod,T,occup,ispinact)}
\begin{tabular}{lcl}
 {\tt ekmod(1:nstate)} & in & given s.p. energies\\
 {\tt T} & in & given temperature\\
 {\tt ispinact} & in & actual spin\\
 {\tt occup(1:kstate)} & in/out & initial occupation and resulting
 Fermi distribution for {\tt T}.\\
\end{tabular}
\\[4pt]
Determines Fermi distribution for given s.p. energies and
temperature. Searches appropriate chemical potential {\tt mu} by
bracketing. Use for repreated calls to {\tt FUNCTION occ}.
\\
\PGRcomm{This routine {\tt fermi\_init} and the related
{\tt FUNCTION occ} are never used, thus obsolete. May be removed.}



\subsubsection*{\tt SUBROUTINE srhomat(psi,aloc,psiorth,occuporth)}
\begin{tabular}{lcl}
 {\tt psi(1:kdfull2,1:kstate)} & in & set of s.p. wavefunctions, not orth-normalized\\
 {\tt psiorth(1:kdfull2,1:kstate)} & out & ortho-normalized natural orbitals\\
 {\tt aloc(1:2*kdfull2)} & in & actual local potential\\
 {\tt occuporth(1:kstate)} & out & occupation numbers for
 ortho-normalized states\\
\end{tabular}
\\[4pt] Computes the density matrix of initial state goiven by set of
wavefunctions {\tt psi} together with their occupations {\tt occup},
the latter communicated through module {\tt params}.  Then
diagonalizes the density matrix and computes on {\tt psiorth} the new
wavefunctions associated with diagonal representation of the density
matrix.  
\\
Finally updates running transformation matrix {\tt psitophi} which is
communicated and stored through module {\tt params}.
\\
\PGRcomm{Usage and propagation of  {\tt psitophi} is somewhat hidden
  because it is handled through a module. Needs to be explained somwhere.}


\subsubsection*{\tt SUBROUTINE scalar(tab1,tab2,scal,ispin, mess)}
\begin{tabular}{lcl}
 {\tt tab1(1:kdfull2,1:kstate)} & in & 1. set of s.p. wavefunctions\\
 {\tt tab2(1:kdfull2,1:kstate)} & in & 2. set of s.p. wavefunctions\\
 {\tt ispin(1:nstate)} & in & spin of s.p. states\\
 {\tt mess} & in & message for print inside routine\\
 {\tt scal(nstate,nstate)} & out & matrix of wavefunction overlaps\\
\end{tabular}
\\[4pt]


\subsubsection*{\tt SUBROUTINE cdiagspin(mat, eigen, vect, N)}
\begin{tabular}{lcl}
 {\tt mat(N,N)} & in & complex Hermitean matrix to be diagonalized\\
 {\tt N} & in & dimension of matrix\\
 {\tt eigen(N)} & out & resulting eigenbvalues\\
 {\tt Vect(N,N)} & out & resulting eigenstates\\
\end{tabular}
\\[4pt]
Driver routine for diagonalization of a complex Hermitean matrix of
dimension {\tt N} which consists in a two blocks for separate spin.
Refers for each single block to routine {\tt cdiag} and subsequent
library routines contained therein.


\subsubsection*{\tt SUBROUTINE indexx (n,arrin,indx)}
\begin{tabular}{lcl}
 {\tt n} & in & length of array\\
 {\tt arrin(1:n)} & in & array to be sorted\\
 {\tt indx(1:n)} & out & pointer array \\
\end{tabular}
\\[4pt]
Evaluates sorting of an array in ascending order.



\subsubsection*{\tt SUBROUTINE occupPerSpin(mess,Occ)}
\begin{tabular}{lcl}
 {\tt mess} & in & character variable with comment printed inside routine\\
 {\tt Occ(1:2)} & out & total number of particles in each spin\\
\end{tabular}
\\[4pt]
Computes number of particles in each sin block. Uses {\tt nstate} and
occupations {\tt occup} from module {\tt params}.



\subsubsection*{\tt CorrectEnergy2(Wref,Eref,w,E,Wout,nloc)}
\begin{tabular}{lcl}
 {\tt W(1:nloc)} & in & initial occupations numbers\\
 {\tt E(1:nloc)} & in & given s.p. energies\\
 {\tt Wref} & in & reference particle number to be reached\\
 {\tt Eref} & in &  reference sum of s.p. energies to be reached\\
 {\tt nloc} & in & actual number of states\\
 {\tt Wout(nloc)} & out  & readjusted occupation numbers\\
\end{tabular}
\\[4pt]
Final energy correction by one step along Fermi distribution
(using Taylor expansion about actual distribution),
see appendix \ref{sec:corriter}).



\subsubsection*{\tt SUBROUTINE ordo\_per\_spin(psi)}
\begin{tabular}{lcl}
 {\tt psi(1:kdfull2,1:kstate)} & in/out & s.p. wavefunctions before
 and after reordering\\
\end{tabular}
\\[4pt]
Reorder states  in two blocks of spin up and down.
Applies that reshuffling to all relevant field of states,
s.p. wavefunctions {\tt psi}, spin per state {\tt ispin},
and occupations {\tt occup}. 
\\
\PGRcomm{Routine has been rendered obsolete by new initialization of
  states which produces immediately the correct sorting. But routine
  should be kept for possible later use (e.g., mixing states from
  different sources.}

\subsubsection*{\tt SUBROUTINE temperature(mu,T)}
\begin{tabular}{lcl}
 {\tt mu} & out & resulting chemical potential\\
 {\tt T} & out & resulting temperature\\
\end{tabular}
\\[4pt]
Takes s.p. energies {\tt amoy} and occupations {\tt occup} from module
{\tt params} and fits a Fermi distribution to it. Temperature and
chemical potentials of the fitted distribution are returned via list.
Calls a fitting routine {\tt lmdif1} using subroutine {\tt ff} as argument.

\subsubsection*{\tt SUBROUTINE ff(m,n,X,FVEC,IFLAG)}
\begin{tabular}{lcl}
 {\tt X(1:n)} & in & array handling chemical potential and temperature\\
 {\tt Fvec(1:m)} & out & array of mismatches of distributions\\
 {\tt n} & in & number of parameters of model, actually 2\\
 {\tt m} & in & number of entries in array\\
 {\tt iflag} & in & flaf possibly written (actually not used)\\
\end{tabular}
\\[4pt]
Mismatch of {\tt occup} (via modules {\tt params}) from Fermi
distribution to given chemical potential and temperature. To be used
in fitting routine {\tt lmdef1}.


\subsubsection*{\tt SUBROUTINE cproject(qin,qout,ispact,q0)}
\begin{tabular}{lcl}
 {\tt qin(1:kdfull2)} & in & s.p. wvaefunction to be projected\\
 {\tt q0(1:kdfull2,1:kstate)} & in & set of s.p. wavefunctions which
 is projected out from {\tt qin}\\
 {\tt ispact} & in & spin associated with {\tt qin}\\
 {\tt qout(1:kdfull2)} & out & projected s.p. wavefunction \\
\end{tabular}
\\[4pt] Projects away from {\tt qin} all contributions of the set 
{\tt q0}.  
\\ 
\PGRcomm{This routine may become obsolete if we recode the
  the variance in routine {\tt calc\_var} to meet the
  standard definition.}


\appendix

\section{The intrinsic excitation energy}
\label{app:eintr}


\PGRcomm{Intrinsic excitation energy yet to be explained. May also
  become part of DCMF explanation.}


\section{Iterative correction of total energy}
\label{sec:corriter}


\PGRcomm{Yet to be imported.}

\bibliographystyle{elsarticle-num}                                                        
\bibliography{relaxtime}             

\end{document}
